---
title: "EDA,recipe and modeltime"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
      smooth_scroll: false
      number_sections: true
      fig_caption: TRUE
    css: "style.css"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
source(here("source","get_lib.R"))
```

<style>
.table-hover > tbody > tr:hover { 
  background-color: #f4f442;
}
</style>

# Get ARTIFACTS 

```{r}
feature_engineering_artifacts_list <- read_rds("model/feature_engineering_artifacts_list.rds")
data_prepared_tbl    <- feature_engineering_artifacts_list$data$data_prepared_tbl
forecast_tbl         <- feature_engineering_artifacts_list$data$forecast_tbl
recipe_spec_2_lag    <- feature_engineering_artifacts_list$recipes$recipe_spec_2
recipe_spec_fourier  <- feature_engineering_artifacts_list$recipes$recipe_spec_fourier 
```

- Splitting `data_prepared_tbl` into `train` and `test` 

```{r}
splits <- data_prepared_tbl %>%
    time_series_split(assess = "8 weeks", cumulative = TRUE)

splits

splits %>%
    tk_time_series_cv_plan() %>%
    plot_time_series_cv_plan(optin_time, optins_trans)
```

# Sequential models

## arima 

- `arima_reg()` is a Modeltime object.
    - All Modeltime algorithms must include a date or date-time feature.
    - This is different than most ML algorithms that don't typically know how to handle dates.
    - Modeltime algorithms use the time-based features to perform calculations like automating period detection and performing feature analysis. 

```{r}
#no XREGS
model_fit_arima <- arima_reg() %>%
    set_engine("auto_arima") %>%
    fit(optins_trans ~ optin_time, data = training(splits))
```


- The author of Modeltime recommend using workflow rather than parsnip object
    - He prefers separating the formula/recipe from the modeling algorithm.
    - it can be convenient to use parsnip fitting when the formula is very basic and not a lot of feature engineering in a recipe. 
    
- `Date Features`
    - These are treated as `non-external regresssors` because the `date-column` follows the target through the entire Modeltime Workflow.
    - Modeltime always has access to the date for every data point in the actual and future data. 
    
- External Regressors
    - Lags & Events are external to core data.
    - Modeltime won't know what these should be unless the you explicitly let the model know
    - Need to create `lag` features outside of preprocessing step using `recipe`

## arima + XREG
    
```{r}
model_spec_arima <- arima_reg() %>%
    set_engine("auto_arima")

workflow_fit_arima <- workflow() %>%
    add_recipe(recipe_spec_fourier) %>%
    add_model(model_spec_arima) %>%
    fit(training(splits))
```

## GLMNET + XREG

```{r}
#elastic net model
model_spec_glmnet <- linear_reg(
    penalty = 0.1,
    mixture = 0.5
) %>%
    set_engine("glmnet")

# #glmnet does not handle by default
# temp <- training(splits)
# temp[!complete.cases(temp), ] %>% tail()
# glmnet_training <- temp[complete.cases(temp), ]

workflow_fit_glmnet <- workflow() %>%
    add_recipe(recipe_spec_2_lag) %>%
    add_model(model_spec_glmnet) %>%
    fit(training(splits))
```


# check point

```{r}
model_tbl <- modeltime_table(
    model_fit_arima,
    workflow_fit_arima,
    workflow_fit_glmnet
)

model_tbl

#update the model description 
model_tbl %>% update_model_description(3, "GLMNET - Lag Recipe")


calibration_tbl <- model_tbl %>%
    modeltime_calibrate(new_data = testing(splits))

calibration_tbl %>%
    modeltime_accuracy()
```

```{r}
?default_forecast_accuracy_metric_set

metric_set(mae, rmse, iic)

calibration_tbl %>%
    modeltime_accuracy(
        metric_set = metric_set(mae, rmse, iic)
    )
```


```{r}
# Table Modeltime Accuracy
calibration_tbl %>%
    modeltime_accuracy(
        metric_set = default_forecast_accuracy_metric_set()
    ) %>%
    table_modeltime_accuracy(
        .interactive = TRUE,
        bordered = TRUE, 
        resizable = TRUE
    )
```

## visualize

```{r}
calibration_tbl %>%
    modeltime_forecast(
        new_data      = testing(splits),
        actual_data   = data_prepared_tbl,
        conf_interval = 0.80
    ) %>%
    plot_modeltime_forecast(
        .legend_max_width = 60,
        .legend_show = FALSE,
        .conf_interval_show = TRUE,
        .conf_interval_alpha = 0.20,
        .conf_interval_fill = "lightblue",
        .title = "Forecast and actual comparision"
    )
```

```{r}
## Refit 
refit_tbl <- calibration_tbl %>%
    modeltime_refit(data = data_prepared_tbl)
```

# Final Forecast

- `new_data` vs `h`

## h

- This parameter is a wrapper for `future_frame` (.length_out = h), where `h` extends the `.calibration_data` `h` steps into the future.
    - see `.calibration_data` only contains time information as feature. 

- important:
    - `h` can only be used if there are no external regressor.
    - Any feature that is not based on the data column is an external regressor.

```{r}
refit_tbl %>%
    modeltime_forecast(
        # h = "8 weeks",
        new_data = forecast_tbl,
        actual_data = data_prepared_tbl,
        conf_interval = 0.80
    ) %>%
    plot_modeltime_forecast(
        .legend_max_width = 25,
        .conf_interval_fill = "lightblue",
        .interactive = TRUE
    ) %>% layout(legend = list(orientation = 'h'))
```

- start from 02_modeltime_new_features.rmd (04)
